---
title: "TAREA 6"
author: "Miguel Gutiérrez"
date: "2025-05-24"
output: html_document
---


## Non-proportional Substitution Patterns I: Generalized Extreme Value Models". 

```{r}
library(mlogit)
library(dfidx)
data("Heating", package = "mlogit")
H <- mlogit.data(Heating, shape = "wide", choice = "depvar", varying = c(3:12))
nl2 <- mlogit(depvar ~ ic + oc, H, nests = list(room = c( 'er', 'gr'), central = c('ec', 'gc', 'hp')), un.nest.el = TRUE, steptol = 1e-12)
```

#### 2. Use model nl2 in this chapter and calculate the direct-point elasticity at the mean values of the variables, for an increase in the installation costs of Gas Central systems.

```{r}
gc_data <- subset(H, alt == "gc")
mean_ic_gc <- mean(gc_data$ic)
probs <- fitted(nl2, type = "probabilities")
mean_prob_gc <- mean(probs[H$alt == "gc"])
beta_ic <- coef(nl2)["ic"]
elasticity_gc_ic <- beta_ic * mean_ic_gc * (1 - mean_prob_gc)
elasticity_gc_ic
```
Un aumento del 1% en el costo de instalación del sistema de Gas Central reduce la probabilidad de su elección en aproximadamente un 0.05213987%.

#### 3. Use model nl2 in this chapter and calculate the cross-point elasticity at the mean values of the variables, for a 1% increase in the operation costs of Gas Central systems.

```{r}
gc_data <- subset(H, alt == "gc")
mean_oc_gc <- mean(gc_data$oc)
probs <- fitted(nl2, type = "probabilities")
mean_probs <- tapply(probs, H$alt, mean)
beta_oc <- coef(nl2)["oc"]
P_ec <- mean_probs["ec"]
P_gc <- mean_probs["gc"]
elasticity_cross_ec_gc <- -beta_oc * mean_oc_gc * (P_gc / P_ec)
elasticity_cross_ec_gc
```
Un incremento del 1% en el costo de operación del sistema de Gas Central aumenta en un 0.03931374% la probabilidad de que los consumidores opten por otra opción

#### 4. Re-estimate the nested logit model in this chapter, but change the nests to types of energy as follows:
 • Gas: gas central, gas room.
 • Electricity: electric central, electric room, heat pump.
Use a single coefficient for the inclusive variables (i.e.,set un.nest.el = TRUE). Are the results reasonable? Discuss.

```{r}
nests_energy <- list(
  gas = c("gc", "gr"),        
  electricity = c("ec", "er", "hp")
)
nl_energy <- mlogit(depvar ~ ic + oc, data = H,
                    nests = nests_energy,
                    un.nest.el = TRUE)
summary(nl_energy)
```
Estos resultados tienen sentido ya que, el mayor costo de instalación y operación del Gas reduce su atractivo, lo cual tiene sentido (mayor precio, menor beneficio para el consumidor).
